---
title: "Thesis - Model"
toc: true
number-sections: true
code-fold: true
warning: false
output: false
error: true
format: gfm
editor_options: 
  markdown: 
    wrap: sentence
---
After failing to come up with meaningful results in model-01 attempt with a linear regression model predicting approved budget, we will now examine the different parameters correlated with varying degrees of SELA budget usage.

# Load libraries

```{r}
library(tidyverse)
library(timeDate)
library(tidymodels)
library(modelr)
library(corrr)
library(ggfortify)
library(DescTools)
library(scales)
library(gtsummary)
library(flextable)
# library(showtext)
```

# Graphics
```{r}
theme_set(theme_light())

# font_add(family = "David", regular = "David.ttf")
# showtext_auto()
# 
# theme_update(
#   text = element_text(family = "David")
# )
```


# Selecting relevant variables
```{r}
sela_df <- df %>% 
  select(
    name,
    muni_id,
    district,
    type,
    pop,
    pop_2015,
    likud_pct,
    coal_pct,
    ses_2013_c,
    peri_2004_c,
    sa_data,
    sector,
    is_nat_pri,
    starts_with("budget")
  )
```

# Checking eligibility for both SELA budget types

## Festival eligibility
The population year determined to use to decide eligibility is 2015. using 2018 created some municipalities that changed population category between the years. the 2015 year was the latest available during the start of 2018.
```{r}
sela_df <- sela_df %>% 
  mutate(
    is_elig_fest = case_when(
      pop_2015 > 100000 ~ FALSE,
      ses_2013_c > 7 ~ FALSE,
      ses_2013_c < 7 ~ TRUE,
      type == "מועצה אזורית" & peri_2004_c <= 2 ~ TRUE,
      is_nat_pri ~ TRUE,
      TRUE ~ FALSE
    ),
    budget_elig_fest = case_when(
      !is_elig_fest ~ 0,
      pop_2015 <= 5000 ~ 70000,
      pop_2015 <= 20000 ~ 115000,
      pop_2015 > 20000 ~ 200000
    )
  )

sela_df %>% 
  ggplot(aes(budget_approved_fest/budget_elig_fest)) +
  geom_histogram()

sela_df %>% 
  filter(!is_elig_fest & budget_approved_fest > 0)

sela_df %>% 
  filter(is_elig_fest & budget_approved_fest < budget_elig_fest)

```
Right now it seems we have two municipalities that are seemingly not eligible for festivals but got budgeted: Hof Ashkelon and Beit Shemesh.

## Initiatives eligibility

The eligibility score of each municipality is calculated: first the unbudgeted municipalities are filtered out and then the score is calculated according the the regulations. the score is normalized between all eligible municipalities and then multiplied by a number close to the total budget allocated. the specific number was decided through trial and error to match the vast majority of municipalities. A ratio between potential budget and score is calculated. Finally, summary charts and tables are presented.
```{r}
sela_df <- sela_df %>% 
  mutate(
    score_elig_init = case_when(
      budget_approved_init == 0 ~ 0,
      pop_2015 <= 10000 ~ 1,
      pop_2015 <= 50000 ~ 2,
      pop_2015 <= 100000 ~ 3,
      pop_2015 <= 150000 ~ 4,
      pop_2015 <= 200000 ~ 5,
      pop_2015 <= 500000 ~ 6,
      pop_2015 > 500000 ~ 7
    ),
    score_elig_init = case_when(
      pop_2015 > 100000 ~ score_elig_init,
      is_nat_pri ~ score_elig_init * 2,
      ses_2013_c <= 6 ~ score_elig_init * 2,
      peri_2004_c <= 2 ~ score_elig_init * 2,
      TRUE ~ score_elig_init
    ),
    budget_elig_init = 23907075 * score_elig_init / sum(score_elig_init)
  )

ratio_elig_init <- sela_df %>% 
  summarize(budget_elig_init / score_elig_init) %>% 
  pull() %>%
  first()

sela_df %>% 
  ggplot(aes(budget_approved_init - budget_elig_init)) +
  geom_histogram()

sela_df %>% 
  count(budget_approved_init - budget_elig_init)

sela_df %>% 
  filter(budget_approved_init > budget_elig_init) %>% 
  arrange(desc(budget_approved_init - budget_elig_init))
```

After calculating the eligibility sum for all municipalities, a further calculation is required for unbudgeted municipalities, as they were filtered out of the previous section. each unbudgeted municipality is calculated a score and a potential budget, multiplied by the ratio calculated in the previous section.
```{r}
sela_df <- sela_df %>% 
  mutate(
    score_elig_init = case_when(
      budget_approved_init > 0 ~ score_elig_init,
      pop_2015 <= 10000 ~ 1,
      pop_2015 <= 50000 ~ 2,
      pop_2015 <= 100000 ~ 3,
      pop_2015 <= 150000 ~ 4,
      pop_2015 <= 200000 ~ 5,
      pop_2015 <= 500000 ~ 6,
      pop_2015 > 500000 ~ 7
    ),
    score_elig_init = case_when(
      budget_approved_init > 0 ~ score_elig_init,
      pop_2015 > 100000 ~ score_elig_init,
      is_nat_pri ~ score_elig_init * 2,
      ses_2013_c <= 6 ~ score_elig_init * 2,
      peri_2004_c <= 2 ~ score_elig_init * 2,
      TRUE ~ score_elig_init
    ),
    budget_elig_init = case_when(
      budget_approved_init > 0 ~ budget_elig_init,
      TRUE ~ score_elig_init * ratio_elig_init
    )
  )

sela_df %>% 
  ggplot(aes(budget_approved_init - budget_elig_init)) +
  geom_histogram()

sela_df %>% 
  count(budget_approved_init - budget_elig_init)

sela_df %>% 
  filter(budget_approved_init > budget_elig_init) %>% 
  arrange(desc(budget_approved_init - budget_elig_init))
```
# Initiatives Execution
Calculating the execution percent of cultural initiatives budget and examining its distribution.
```{r}
sela_df <- sela_df %>% 
  mutate(
    exec_pct_init = budget_approved_init / budget_elig_init
  )

sela_df %>% 
  ggplot(aes(exec_pct_init)) +
  geom_histogram()
```
Visualizing the relationships between relevant variables and the execution percent. Box plots for categorical variables and scatter plots for continuous variables.
```{r}
sela_df %>% 
  filter(budget_elig_init - budget_approved_init > 1000) %>% 
  select(
    name,
    district,
    type,
    ses_2013_c,
    peri_2004_c,
    sector,
    exec_pct_init
  ) %>% 
  mutate(across(-exec_pct_init, ~ as_factor(.))) %>% 
  pivot_longer(!c(name, exec_pct_init), names_to = "var", values_to = "value") %>% 
  ggplot(aes(value, exec_pct_init)) + 
  geom_boxplot() + 
  facet_wrap(~ var, scales = "free_y") +
  coord_flip()

sela_df %>% 
  filter(budget_elig_init - budget_approved_init > 1000) %>% 
  select(
    name,
    pop_2015,
    likud_pct,
    coal_pct,
    exec_pct_init
  ) %>% 
  mutate(pop_2015_log10 = log10(pop_2015)) %>% 
  pivot_longer(!c(name, exec_pct_init), names_to = "var", values_to = "value") %>% 
  ggplot(aes(value, exec_pct_init)) + 
  geom_point() +
  facet_wrap(~ var, scales = "free_x") +
  geom_smooth(se = FALSE)

sela_df %>% 
  filter(budget_elig_init - budget_approved_init > 1000) %>% 
  summarise(cor(log10(pop_2015), exec_pct_init))
```
# Eligibilty linear modelling
## Per capita modelling
### Transform dependant variable
```{r}
sela_df <- sela_df %>% 
  mutate(
    budget_elig_tot = budget_elig_fest + budget_elig_init,
    budget_elig_tot_capita = budget_elig_tot / pop_2015
  )

sela_df %>% 
  ggplot(aes(budget_elig_tot_capita)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density()

sela_df %>% 
  ggplot(aes(log10(budget_elig_tot_capita))) +
  geom_histogram(aes(y = ..density..)) +
  geom_density()

sela_df %>% 
  ggplot(aes(sqrt(budget_elig_tot_capita))) +
  geom_histogram(aes(y = ..density..)) +
  geom_density()

sela_df %>% 
  ggplot(aes((budget_elig_tot_capita)^0.3)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density()

sela_df %>% 
  transmute(
    budget_elig_tot_capita_log10 = log10(budget_elig_tot_capita),
    budget_elig_tot_capita_pw.1 = budget_elig_tot_capita ^ 0.1,
    budget_elig_tot_capita_pw.2 = budget_elig_tot_capita ^ 0.2,
    budget_elig_tot_capita_pw.3 = budget_elig_tot_capita ^ 0.3,
    budget_elig_tot_capita_pw.4 = budget_elig_tot_capita ^ 0.4,
    budget_elig_tot_capita_pw.5 = budget_elig_tot_capita ^ 0.5
  ) %>% 
  pivot_longer(everything(), names_to = "fn", names_prefix = "budget_elig_tot_capita_", values_to = "value") %>% 
  group_by(fn) %>% 
  summarise(
    skewness = skewness(value),
    kurtosis = kurtosis(value)
  )
```
It seems that the transformation of per capita to the power of 0.3 creates the most normally distributed variable with the least skewness and kurtosis.

### Linear model per capita transformed
```{r}
sela_df <- sela_df %>% 
  mutate(budget_elig_tot_capita_pw.3 = budget_elig_tot_capita ^ 0.3)

sela_mdl1 <- lm(budget_elig_tot_capita_pw.3 ~ sector * likud_pct, data = sela_df)

sela_mdl1 %>% 
  tidy()

sela_mdl1 %>% 
  glance()

sela_mdl1 %>% 
  augment() %>% 
  ggplot(aes(likud_pct, color = sector)) + 
  geom_point(aes(y = budget_elig_tot_capita_pw.3)) +
  geom_line(aes(y = .fitted))
```
### Linear model per capita untransformed
```{r}
sela_mdl1a <- lm(budget_elig_tot_capita ~ sector * likud_pct, data = sela_df)

sela_mdl1a %>% 
  tidy()

sela_mdl1a %>% 
  glance()

sela_mdl1a %>% 
  augment() %>% 
  ggplot(aes(likud_pct, budget_elig_tot_capita, color = sector)) + 
  geom_line(aes(y = .fitted)) +
  geom_point()
```
The per capita budget modelling, both transformed and untransformed, went unsuccessful, as the slope coefficient for the jewish sector was negative.
We will now examine total budget rather than per capita budget.

## total budget modelling
### Check distribution of dependant variable
```{r}
sela_df %>% 
  ggplot(aes(budget_elig_tot)) +
  geom_histogram() +
  theme_minimal() +
  scale_x_continuous(labels = label_comma()) +
  labs(
    x = 'גובה הזכאות של רשות מקומית לתמיכה תקציבית של תקנת סל"ע בשנת 2018 (ש"ח)',
    y = "מספר רשויות"
  )
  

sela_df %>% 
  summarise(
    mean = mean(budget_elig_tot),
    sd = sd(budget_elig_tot),
    skewness = skewness(budget_elig_tot),
    kurtosis = kurtosis(budget_elig_tot)
  )
```

### Distribution table of all variables
```{r}
sela_df %>% 
  select(
    budget_elig_tot,
    pop_2015,
    ses_2013_c,
    peri_2004_c,
    sector,
    likud_pct
  ) %>% 
  tbl_summary(
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  )
```


### Model with Likud voting percent
```{r}
sela_mdl2 <- lm(budget_elig_tot ~ sector * likud_pct, data = sela_df)

sela_mdl2 %>% 
  tidy()

sela_mdl2 %>% 
  glance()

sela_mdl2 %>% 
  augment() %>% 
  ggplot(aes(likud_pct, budget_elig_tot, color = sector)) + 
  geom_line(aes(y = .fitted)) +
  geom_point()
```
This shows a good relationship so that jewish municipalities are receiveing more budget as the Likud voting percent rises.

### Model with coalition voting percent
```{r}
sela_mdl2 <- lm(budget_elig_tot ~ sector * coal_pct, data = sela_df)

sela_mdl2 %>% 
  tidy()

sela_mdl2 %>% 
  glance()

sela_mdl2 %>% 
  augment() %>% 
  ggplot(aes(coal_pct, budget_elig_tot, color = sector)) + 
  geom_line(aes(y = .fitted)) +
  geom_point()
```
This model is even stronger, also showing a positive relationship between total SELA budget and coalition parties voting percentage, but with an adjusted R-Squared of 0.22.

Let's check if it still stands after adding control variables

```{r}
sela_mdl3 <- lm(budget_elig_tot ~ sector : likud_pct + pop_2015 + ses_2013_c + peri_2004_c, data = sela_df)
sela_mdl3_control <- lm(budget_elig_tot ~ pop_2015 + ses_2013_c + peri_2004_c, data = sela_df)
sela_mdl3_control2 <- lm(budget_elig_tot ~ pop_2015 + ses_2013_c + peri_2004_c + is_nat_pri, data = sela_df)
anova(sela_mdl3_control, sela_mdl3)

sela_mdl3 %>% 
  tidy()

sela_mdl3 %>% 
  glance()

sela_mdl3_control %>% 
  tidy()

sela_mdl3_control %>% 
  glance()

sela_mdl3 %>% 
  augment() %>% 
  ggplot(aes(likud_pct, budget_elig_tot, color = sector)) + 
  geom_line(aes(y = .fitted)) +
  geom_point()

sela_mdl3 %>% 
  augment() %>% 
  ggplot(aes(likud_pct, .fitted - sela_mdl3_control %>% augment() %>% pull(.fitted), color = sector)) + 
  geom_line() +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
The model is still significant for Likud voting percentage after adding control variables.
This is not the case for coalition voting percentage.

Let's create a better plot for the last plot of the difference between the two models.
```{r}
sela_mdl3 %>% 
  augment() %>% 
  ggplot(aes(
    likud_pct, .fitted - sela_mdl3_control %>% augment() %>% pull(.fitted),
    color = sector,
    shape = sector
  )) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(labels = label_comma()) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_color_discrete(labels = c("ערבי", "יהודי")) +
  scale_shape_discrete(labels = c("ערבי", "יהודי")) +
  guides(shape = guide_legend(override.aes = list(size = 3))) +
  theme(legend.position = "bottom") +
  labs(
    x = "אחוז הצבעה לליכוד בבחירות 2015",
    y = '\u202B"דיבידנד הנאמנות בתרבות" (ש"ח)',
    color = "מגזר",
    shape = "מגזר"
  )

```


Let's put the two models in a table.

```{r}
tbl_merge(
  list(
    tbl_regression(
      sela_mdl3_control,
      intercept = TRUE,
      conf.int = FALSE
    ) %>% 
      modify_header(
        statistic = "**t-statistic**",
        std.error = "**SE**"
      ) %>% 
      bold_p() %>% 
      add_glance_table(
        include = c(r.squared, statistic)
      )
    ,
    tbl_regression(
      sela_mdl3,
      intercept = TRUE,
      conf.int = FALSE
    ) %>% 
      modify_header(
        statistic = "**t-statistic**",
        std.error = "**SE**"
      ) %>% 
      bold_p() %>% 
      add_glance_table(
        include = c(r.squared, statistic)
      )
  ),
  tab_spanner = c("M1", "M2")
) %>% 
  modify_table_body(~.x %>% arrange(row_type == "glance_statistic")) %>% 
  as_flex_table()
```

Let's check the distribution of voting to Likud across SES clusters.
```{r}
elec_more_df <- df %>% 
  mutate(
    likud_votes = good_votes * (likud_pct / 100),
    ses_2013_c = factor(ses_2013_c)
  )

elec_more_df %>% 
  group_by(ses_2013_c) %>% 
  summarise(likud_pct = round(sum(likud_votes) / sum(good_votes) * 100, 1)) %>% 
  ggplot(aes(ses_2013_c, likud_pct)) +
  geom_col() + 
  geom_text(
      aes(label = paste0(round(likud_pct, digits = 1), "%")),
      vjust = -0.5
  ) +
  scale_y_continuous(
    labels = label_percent(scale = 1),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    x = "\u202Bאשכול חברתי-כלכלי (2013)",
    y = "\u202Bאחוז הצבעה לליכוד בבחירות 2015"
  )
```


Since SES cluster 7 had special criteria, Let's look at the Likud voting patterns within them.

```{r}
ses_7_df <- elec_more_df %>% 
  filter(ses_2013_c == "7")

ses_7_df %>% 
  count(is_nat_pri, type == "מועצה אזורית" & peri_2004_c <= 2)

ses_7_df %>% 
  group_by(is_nat_pri) %>% 
  summarise(likud_pct = round(sum(likud_votes) / sum(good_votes) * 100, 1)) %>% 
  ggplot(aes(is_nat_pri, likud_pct)) +
  geom_col() + 
  geom_text(
      aes(label = paste0(round(likud_pct, digits = 1), "%")),
      vjust = -0.5
  ) +
  scale_y_continuous(
    labels = label_percent(scale = 1),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    x = "אזור עדיפות לאומית",
    y = "אחוז הצבעה לליכוד"
  )

# by a simple mean and not total mean
ses_7_df %>% 
  group_by(is_nat_pri) %>% 
  summarise(likud_pct = round(mean(likud_pct), 1)) %>% 
  ggplot(aes(is_nat_pri, likud_pct)) +
  geom_col() + 
  geom_text(
      aes(label = paste0(round(likud_pct, digits = 1), "%")),
      vjust = -0.5
  ) +
  scale_y_continuous(
    labels = label_percent(scale = 1),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    x = "אזור עדיפות לאומית",
    y = "אחוז הצבעה לליכוד"
  )

# by eligibility criterea for initiatives
elec_more_df %>% 
  filter(ses_2013_c %in% c("7", "8", "9", "10")) %>% 
  mutate(special_elig_init = is_nat_pri | peri_2004_c <= 2) %>% 
  group_by(special_elig_init) %>% 
  summarise(likud_pct = round(mean(likud_pct), 1)) %>% 
  ggplot(aes(special_elig_init, likud_pct)) +
  geom_col() + 
  geom_text(
      aes(label = paste0(round(likud_pct, digits = 1), "%")),
      vjust = -0.5
  ) +
  scale_y_continuous(
    labels = label_percent(scale = 1),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    x = "אזור עדיפות לאומית או רשות פריפריאלית",
    y = "אחוז הצבעה לליכוד"
  )
```



The next thing to look at is the difference between the eligibility for SELA budget of each municipality and the hypothetical budget it would have received according to the 2014 distribution of total Ministry of Culture budget to organizations by their municipality.

## difference in hypothetical distribution modelling
### Create the difference
```{r}
sela_df <- sela_df %>% 
  mutate(
    budget_pct_tot_2014 = budget_approved_2014 / sum(budget_approved_2014), # percent of total 2014 culture budget going to this municipality
    budget_sela_hypo_2018 = sum(budget_elig_tot) * budget_pct_tot_2014, # the hypothetical SELA budget that would be going to the municipality in 2018 if the budget was distributed by the 2014 distribution
    budget_sela_hypo_diff = budget_elig_tot - budget_sela_hypo_2018, # the difference between the real SELA budget eligibility of a municipality in 2018 and the hypothetical eligibility by 2014 distribution
    budget_sela_hypo_diff_capita = budget_sela_hypo_diff / pop_2015
  )

sela_df %>% 
  filter(budget_sela_hypo_diff > -500000) %>% 
  ggplot(aes(budget_sela_hypo_diff)) +
  geom_histogram()

sela_df %>% 
  filter(budget_sela_hypo_diff < -500000) %>% arrange(budget_sela_hypo_diff)

sela_df %>% 
  ggplot(aes(budget_sela_hypo_diff_capita)) +
  geom_histogram() + 
  geom_density(aes(y = ..count..))

sela_df %>% 
  filter(budget_sela_hypo_diff < -500000) %>% arrange(budget_sela_hypo_diff)

sela_df %>% 
  mutate(budget_sela_hypo_diff_log10 = log10(1 - min(budget_sela_hypo_diff) + budget_sela_hypo_diff)) %>% 
  ggplot(aes(budget_sela_hypo_diff_log10)) +
  geom_histogram()
```
### Model the hypothetical difference per capita
```{r}
sela_df %>% 
  select(-c(name, muni_id, budget_sela_hypo_diff_capita, sa_data, contains("budget"))) %>% 
  map(~ lm(budget_sela_hypo_diff_capita ~ .x, data = sela_df )) %>% 
  map(glance) %>% 
  map("r.squared") %>% 
  as_tibble() %>% 
  pivot_longer(everything() ,names_to = "var", values_to = "r.squared") %>% 
  arrange(desc(r.squared))
  
lm(budget_sela_hypo_diff_capita ~ peri_2004_c + ses_2013_c + sector * coal_pct + type , data = sela_df) %>% 
  glance()
lm(budget_sela_hypo_diff_capita ~ peri_2004_c + ses_2013_c + sector * coal_pct + type , data = sela_df) %>% 
  tidy()
```

Can't model the hypothetical budget with political variables.

# Gini

```{r}
df %>% 
  summarise(
    gini_2018 = Gini(budget_approved, weights = pop),
    gini_2014 = Gini(budget_approved_2014, weights = pop)
  )
```

