---
title: "Thesis - Model"
toc: true
number-sections: true
code-fold: true
warning: false
output: false
error: true
format: gfm
editor_options: 
  markdown: 
    wrap: sentence
---
After failing to come up with meaningful results in model-01 attempt with a linear regression model predicting approved budget, we will now examine the different parameters correlated with varying degrees of SELA budget usage.

# Load libraries

```{r}
library(tidyverse)
library(tidymodels)
library(modelr)
library(corrr)
library(ggfortify)
```

# Selecting relevant variables
```{r}
sela_df <- df %>% 
  select(
    name,
    muni_id,
    district,
    type,
    pop,
    likud_pct,
    coal_pct,
    ses_2013_c,
    peri_2004_c,
    sa_data,
    sector,
    is_nat_pri,
    starts_with("budget")
  )
```

# Checking eligibility for both SELA budget types

## Add variables for eligibilty
```{r}
sela_df <- sela_df %>% 
  mutate(
    is_elig_fest = case_when(
      pop > 100000 ~ FALSE,
      ses_2013_c > 7 ~ FALSE,
      ses_2013_c < 7 ~ TRUE,
      type == "מועצה אזורית" & peri_2004_c <= 2 ~ TRUE,
      is_nat_pri ~ TRUE,
      TRUE ~ FALSE
    )
  )

sela_df %>% 
  group_by(is_elig_fest) %>% 
  summarise(sum = sum(budget_approved_fest))

check <- sela_df %>% 
  filter(!is_elig_fest & budget_approved_fest > 0)
```
Right now it seems we have two municipalities that are seemingly not eligible but got budgeted: Hof Ashkelon and Beit Shemesh.
